Elephant Seal Age JAE Intraspecific Variation

### 2023-12-06 paper/paperv3.qmd
fixing the average line on phenology plot

```{r}
#| label: fig-repro_age
#| fig-cap: "Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age."
#| fig-width: 9

@@

surv_recruit_summary %>% 
  ggplot(aes(x = age, y = mean, color = metric, fill = metric)) + 
  geom_ribbon(aes(ymin = mean_lwr, ymax = mean_upr, alpha = metric),
              surv_recruit_pred,
              color = NA) +
  geom_pointrange(aes(ymin = low, ymax = high)) +
  geom_line(aes(linetype = metric),
            surv_recruit_pred) +
  scale_y_continuous(labels = scales::percent) + 
  scale_color_manual(values = c("#829BD1", "#E9B668")) + 
  scale_fill_manual(values = c("#829BD1", "#E9B668")) + 
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_alpha_manual(values = c(0.25, 0.1)) +
  expand_limits(y = c(0, 0.5)) +
  labs(x = "Maternal Age (Years)", y = "Pup Survival or Recruitment/n") +
  theme_few() + 
  theme(legend.position = c(0.1,0.97), 
        legend.justification = c(0,1), 
        legend.title = element_blank(), 
        legend.direction = "horizontal", 
        axis.title = element_text(size = 13))

@@

year_ranef <- ranef(pup_survival_model)$year %>% 
  as_tibble(rownames = "year") %>% 
  mutate(year = as.numeric(year)) %>% 
  rename(intercept = `(Intercept)`)

@@

```{r}
#| label: fig-sex_ratio
#| fig-cap: "Pup sex skews slightly towards males with increasing maternal age, but it's not significant."
#| fig-width: 9
#A: Pup sex shifts towards males with increasing maternal age, but it's not significant.

@@

mean_phenology <- seal_phenology %>% 
  summarize(across(BA:MD, mean), 
            .groups = "drop") %>% 
  pivot_longer(everything(),
               names_to = "event", 
               values_to = "doy")

@@

The table is supplemental.

```{r}
#| label: fig-phenology
#| fig-height: 7
#| fig-width: 7
#| fig-cap: "Successful breeding and increased maternal age both impact elephant seal annual cycles. A: Visualization of life history timing across ages, faceted by breeding status. White dashed lines on each panel represent the raw data from the other panel to facilitate comparisons between breeders and non-breeders. B: LMM coefficients for phenological parameters. The units for maternal age coefficients are in days per year. The coefficients for breeding status are in days. Black points are significant, gray points are not significant."

@@

ggsave("outputs/agefig4.png", plot = , height = 9, width = 9)

table(seal_phenology$age,seal_phenology$observed)

#average the averages - global average of age-specific averages 
```

Supplemental table

phenology_models %>% 
  select(-c(model_data, model)) %>% 
  pivot_longer(age_coef:breeder_signif, 
               names_to = c("param", "metric"),
               names_sep = "_",
               values_to = "value") %>% 
  pivot_wider(names_from = metric, values_from = value) %>% 
  mutate(coef = str_glue("{round(coef, 2)} +- {round(stderr, 2)}"),
         signif = str_glue("p = {round(pval, 4)} ({signif == 1})")) %>% 
  select(-c(stderr, pval)) %>% 
  pivot_wider(names_from = param, values_from = c(coef, signif)) %>% 
  select(phenology, ends_with("age"), ends_with("breeder")) %>% 
  knitr::kable(col.names = c("Phenology", 
                             "Age coef. (days/year)", 
                             "Significance", 
                             "Breeding status coef. (days)", 
                             "Significance")) %>% 
  kableExtra::kable_styling()

>>>>

```{r}
#| label: fig-repro_age
#| fig-cap: "Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age."
#| fig-width: 4.5

@@

surv_recruit_summary %>% 
  ggplot(aes(x = age, y = mean, color = metric, fill = metric)) + 
  geom_ribbon(aes(ymin = mean_lwr, ymax = mean_upr, alpha = metric),
              surv_recruit_pred,
              color = NA) +
  geom_pointrange(aes(ymin = low, ymax = high)) +
  geom_line(aes(linetype = metric),
            surv_recruit_pred) +
  scale_y_continuous(labels = scales::percent) + 
  scale_color_manual(values = c("#829BD1", "#E9B668")) + 
  scale_fill_manual(values = c("#829BD1", "#E9B668")) + 
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_alpha_manual(values = c(0.25, 0.1)) +
  expand_limits(y = c(0, 0.5)) +
  labs(x = "Maternal Age (Years)", y = "Pup Survival or Recruitment") +
  theme_few() + 
  theme(legend.position = c(0.1,0.97), 
        legend.justification = c(0,1), 
        legend.title = element_blank(), 
        legend.direction = "horizontal", 
        axis.title = element_text(size = 13))

@@

year_ranef <- ranef(pup_surv_model)$year %>% 
  as_tibble(rownames = "year") %>% 
  mutate(year = as.numeric(year)) %>% 
  rename(intercept = `(Intercept)`)

@@

```{r}
#| label: fig-sex_ratio
#| fig-cap: "Pup sex skews slightly towards males with increasing maternal age, but it's not significant."
#| fig-width: 4.5
#A: Pup sex shifts towards males with increasing maternal age, but it's not significant.

@@

mean_phenology <- seal_phenology %>% 
  group_by(age) %>% 
  summarize(across(BA:MD, mean)) %>% 
  summarize(across(BA:MD, mean)) %>% 
  pivot_longer(everything(),
               names_to = "event", 
               values_to = "doy")

@@

```{r}
#| label: fig-phenology
#| fig-height: 4.5
#| fig-width: 7
#| fig-cap: "Successful breeding and increased maternal age both impact elephant seal annual cycles. A: Visualization of life history timing across ages, faceted by breeding status. White dashed lines on each panel represent the raw data from the other panel to facilitate comparisons between breeders and non-breeders. B: LMM coefficients for phenological parameters. The units for maternal age coefficients are in days per year. The coefficients for breeding status are in days. Black points are significant, gray points are not significant."

@@

# table(seal_phenology$age,seal_phenology$observed)

#white dashed lines average the averages - global average of age-specific averages 
```
